Who introduced this bug? \\ It may not have been caused by the previous commit!

### 2016-02-01 paperMSR.tex
some changes

For that, we had to analyze the lines involved in the bug fix and in the \emph{parent} commit of the bug fix commit, being sure that the lines were added, inserted or modified in the previous commit. We refer to \emph{parent} commit as the commit that modified any line of code in the file before the fix-bug commit, in contrast to the \emph{previous} commit where the modified lines were the same than in the fix-bug commit. It should be noted that those lines modified in the parent commit do not have to be the ones that have been modified in the bug-fix.
Figure~\ref{fig:parentgerrit} contains a snapshot of the information provided by Gerrit, where the link to the parent commit(s) can be found, that corresponds to the bug-fix shown in~\ref{fig:1}. As can be seen, the previous commit (\textit{31f08423}) in Figure~\ref{fig:1} is different from the parent commit displayed in Figure~\ref{fig:parentgerrit} (\textit{db7fc59ebc}).

We do this process to be sure that we are looking the correct change, because sometimes although the commit added many lines, if you look the code before the commit you can check that some of the lines added was there, and in that case, it is a false positive where the previous commit did not cause the bug.

%\grex{Gema, puedes aclarar este parrafo? Que es el "parent" commit? Es lo mismo que el previous commit? Es la primera vez que hablamos de el. Quizas sea buena idea que lo reformules en castellano, porque ahora mismo es criptico}\gema{el parent commit es el ultimo commit que se hizo en el fichero antes de que se realizase el que arreglo el error, por tanto lo que hago usar diff y blame en el commit parent y en el commit que arregla el error para saber que lineas se han cambiado y que commit las cambio), y despues hago lo mismo con el commit previo y el commit parent del previo para saber que en realidad en ese momento se produjo el cambio. En el gerrit nos dan esa info como se puede ver en la figura}

%\grex{Sigue sin quedarme clara cual es la relacion entre el parent commit y el previous commit (mira que yo pensaba que eran lo mismo)... podriamos dar una definicion?}  
%\gema{EL parent commit es el ultimo commit que toco alguna linea del fichero antes de que se realizase el fix bug commit, pero las lineas que toco el commit parent no tienen porque contener ningun bug. Mientras que el commit previo es la ultima modificacion que se hizo en aquellas lineas que ha tocado el bug-fix.Estaría bien dar una definicion por si no se ve claro}

\begin{figure}[ht]
\centering
\includegraphics[height=2cm]{parent.png}
\caption{Information about the bug as displayed by Gerrit.}
\label{fig:parentgerrit}       % Give a unique label
\end{figure}

@@

We classify a total of 459 tickets using the tool, resulting in 917 reviews\grex{Mirando la Tabla 2, veo que las revisiones suman 917, no 918!} \gema{Es cierto, son 917}. Only those tickets classified as bug reports by both researchers were considered in the next stage, which analyzes if the cause of the bug was introduced in their previous commits. This process requires manual inspection by researchers. In the mean classifying a ticket takes \gema{between 5 and 10 minutes per ticket}, although the amount of time decreases with experience as could be expected.

@@

In this stage we analyzed the 189 tickets classified \grex{Pero no eran 209?} \gema{deberian ser 209 pero me han quedado 20 bugs report por analizar, si tengo tiempo cuando llegue a casa lo hago y anado los numeros} \grex{no se si hace falta, a ver que dice jesus} as bug reports and their previous commits.
The possible outcome of the analysis was one of the following three options:

@@

This analysis takes into account that the bug could span many lines that may belong to several previous commits, but in fact, not all of them could had caused the bug. It may happen that in the previous commit, lines have been copied from further previous commits, comments may have been modified or blank spaces/lines have been introduced. Hence, the cause could be found in a single previous commit, in many or even in none.

Figure~\ref{fig:manyprevious} contains a real example of a previous commit where more than one commit has been identified. In this case, we have two possible commits: \textit{e7be0a988} and \textit{e5296c1da}. Previous commit \textit{e7be0a988} did not cause the bug, because the modification affects only the version number of the software. It is previous commit \textit{e5296c1da} the one that caused the bug, because it introduced an incorrect \texttt{break} line. %\grex{Todavia no esta del todo claro... yo creo que esto merece un ejemplo - incluso con codigo real, tal y como hemos hecho al principio del articulo. Ademas, nos lo podemos permitir; tenemos espacio.}\gema{perfecto, ahora busco un ejemplo y lo anado abajo}

\grex{No estan mal las descripciones superiores en la figura~\ref{fig:manyprevious} que dice ``After Fix Bug'' y ``Fix-inducing (Before fix bug)''?}

\begin{figure}[ht]
\centering
\includegraphics[height=2.5cm]{manyprevious.png}
\caption{Previous commits: \textit{e7be0a988} (left) and \textit{e5296c1da} (right).}
\label{fig:manyprevious}       % Give a unique label
\end{figure}

@@

As can be seen in Table~\ref{tab:responsability}, from the 308 previous commits, 152 have been considered to be the cause of the bug, whereas 114 were identified as not beign the cause. We were unable to decide in 42 cases, as only lines had been added or due to our limited knowleadge about the code. \grex{Gema, puedes poner un ejemplo de estas dos cosas? Uno de solo added lines y otro de limited knowledge. Si el ejemplo es verdadero, mejor.}\gema{perfecto, ahora busco un ejemplo y lo anado abajo}
We discarded 40 more because they were false positives (\emph{noise}) such as blank lines, changes in comments or even a change in the version of the file.

\begin{table}[htb]
\begin{center}
\begin{tabular}{lrr}
\toprule[0.3mm]
  & \multicolumn{1}{c}{Before} & \multicolumn{1}{c}{After} \\
  & \multicolumn{1}{c}{Deleting Noise} & \multicolumn{1}{c}{Deleting Noise} \\\hline
{Cause} & (152) 44\% & (152) 49\% \\[0ex]
{Not Cause} & (154) 44\% & (114) 37\% \\[0ex]
{Undecided} & (42) 12\% & (42) 14\% \\[0ex]
\bottomrule[0.3mm]
\end{tabular}
\caption{Number of times (and percentage) where the previous commit is the cause, not the cause or could not be classified, before and after deleting noise.}
\label{tab:responsability}
\end{center}
\end{table}

@@

We also studied the distribution of the number of previous commits for each bug. This result will provide further insight into the bug-seeding nature; it offers as well an idea of the complexity of identifying the cause of a bug, as the more commits involved, the harder it is to identify the cause and understand it. As shown in Table~\ref{tab:secondStage2}, usually the number of commits that can be considered as previous is 1 (over 69\% of the cases), followed by 2 commits (13\%). In around 10\% of the cases, 3 or more commits are involved.

% the number of commits that can be considered as previous is 1, followed by 2 commits. \grex{the number of commits that can be considered as previous is 1, followed by 2 commits. No se que queremos decir} \grex{Tenemos que responder a estas dos preguntas todavia: Por que es interesante hacer lo que se comenta en este parrafo? Que utilidad tiene?}\gema{Creo que era interesante saber como se distribuyen los causantes de los bugs en cada projecto y cuantos commits previos presentan los bugs report en cada projecto analizado, ademas cuantos mas commits previos, en ocasiones, es muy dificil saber quien es el responsable por tanto pense que podria haber algun patron en relacion con el numero de commits previos y cuantos de ellos son responsables, pero parece que esto no ocurre con la muestra que hemos analizado.}

\begin{table*}[htb]
\begin{center} {\footnotesize
\begin{tabular}{lccccc}
\toprule[0.3mm]
  & \multicolumn{1}{c}{One previous} & \multicolumn{1}{c}{two previous} & \multicolumn{1}{c}{three previous} & \multicolumn{1}{c}{four previous} & \multicolumn{1}{c}{+five previous}\\
  & \multicolumn{1}{c}{commit} & \multicolumn{1}{c}{commit} & \multicolumn{1}{c}{commit}& \multicolumn{1}{c}{commit}& \multicolumn{1}{c}{commit}\\\hline
\raisebox{1ex}{Neutron} & 11 & 3 & 2 & 2 & 0 \\[0ex]
\raisebox{1ex}{Horizon} & 39 & 8 & 3 & 2 & 4 \\[0ex]
\raisebox{1ex}{Nova} & 44 & 5 & 2 & 4 & 4 \\[0ex]
\raisebox{1ex}{Cinder} & 37 & 9 & 6 & 2 & 2\\[0ex]
\raisebox{1ex}{Total} & 131 & 25 & 13 & 10 & 10\\[0ex]
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Distribution of the number of commits that can be considered as the previous commit per bug report for each project.}
\label{tab:secondStage2}
\end{center}
\end{table*}

@@

In any case, our research shows evidence that assuming that the previous commit is where the cause of a bug can be found does not hold for a not insignificant percentage of bugs. \grex{Tengo que elaborar mas esto, que es el punto mas impactante del articulo.}

@@

As any other empirical study, this one presents several threats to its validity, external and internal, that have to be considered and taken into account. In order to allow others to study it in detail, replicate it or even build on top of it, we have set up a replication package\footnote{\url{http://gsyc.urjc.es/~grex/repro/2016-msr-prevcommit}} including data sources, intermediate data and scripts that can be obtained.

@@

\begin{itemize}
  \item We have not considered those tickets where the two researchers showed discordance. 
%  \begin{enumerate}
%    \item Should they discuss after their analysis to reach a better classification?, Should the tool provide this?
%    \item Does the Bug report only the same ticket classified as Bug report for all the researchers?
%  \end{enumerate}
    \item We have not taken into account errors that have been classified into \textit{Undecided}, and probably we have lost some \emph{actual} bug reports.
    \item There could be some lax criteria involving the subjective opinion of the researchers.
    \item Although the researchers are experienced programmers, they are not experts in the  OpenStack project, and their inexperience may have influenced the results of the analysis.
    \item We are only using part of the information that the tickets provide, like comments and text. There could be a recognized pattern in the data, unknown at first sight, that involves other parts of the information.
    \item We have used a random script to extract the tickets from Launchpad that have been reported during 2015. There could be unintended bias of the data, because many reasons, as for instnace the phase of the project.
    \item In some cases, researchers may have classified the previous commit as the cause of the bug, even if this may not be the case (see discussion on additional conditions in if statements).
\end{itemize}

@@

With our methodology we have identifie which are real changes that introduced the bug, and this could be useful to improve the accuracy of those tools developed to prevent bugs. Also, the software developers stand to benefit from identifying where the bug was inserted, improving their methodology.
%How to plan to apply the knowledge in a way that can benefit software engineers.
%What are the anticipated contributions of the work?  How will you evaluate them to demonstrate usefulness?

A final field of future work could be concerned with the full automation of the methodology, developing  an automatic classifier based on the idea that not all the previous commits inject the bug. \grex{mejorar este parrafo.}

Another interesting investigation could perform the same empirical study on a project with a less active community, to prove if our idea is fulfill in other projects. So, for projects that do not evolve as rapidly as OpenStack, e.g., those with a more stable API, could offer completely different results even if applying the same analysis methodology used in this paper.

>>>>

For that, we had to analyze the lines involved in the bug fix commit and in its \emph{parent} commit, being sure that the lines were added, inserted or modified in the previous commit. We refer to \emph{parent} commit as the last commit that was edited to produce the subsequent state of the file before the bug fix and could be different from the commit previous. We do this process to be sure that we are looking the correct change, because some times although the commit added many lines, if you look the code before the commit you can check that some of the lines added was there, and in that case, it is a false positive where the previous commit did not cause the bug. \grex{Gema, puedes aclarar este parrafo? Que es el "parent" commit? Es lo mismo que el previous commit? Es la primera vez que hablamos de el. Quizas sea buena idea que lo reformules en castellano, porque ahora mismo es criptico}\gema{el parent commit es el ultimo commit que se hizo en el fichero antes de que se realizase el que arreglo el error, por tanto lo que hago usar diff y blame en el commit parent y en el commit que arregla el error para saber que lineas se han cambiado y que commit las cambio), y despues hago lo mismo con el commit previo y el commit parent del previo para saber que en realidad en ese momento se produjo el cambio. En el gerrit nos dan esa info como se puede ver en la figura~\ref{fig:parent}}

\grex{Sigue sin quedarme clara cual es la relacion entre el parent commit y el previous commit (mira que yo pensaba que eran lo mismo)... podriamos dar una definicion?}  
\gema{EL parent commit es el ultimo commit que toco alguna linea del fichero antes de que se realizase el fix bug commit, pero las lineas que toco el commit parent no tienen porque contener ningun bug. Mientras que el commit previo es la ultima modificacion que se hizo en aquellas lineas que ha tocado el bug-fix.Estaría bien dar una definicion por si no se ve claro}
The information about the bug fix commit is shown in the figure \ref{fig:6}, this information has been extracted from the gerrit of the figure \ref{fig:1} and we can apreciate that the previous commit in the example \ref{fig:1}(\textit{31f08423}) is different from the parent commit displayed in the figure \ref{fig:6},(\textit{db7fc59ebc}).
\begin{figure}[ht]
\centering
\includegraphics[height=2cm]{parent.png}
\label{fig:6}       % Give a unique label
\caption{Information about the bug fix commit displayed in gerrit. }
\end{figure}

@@

We classify a total of 459 tickets using the tool, resulting in 917 reviews. Only those tickets classified as bug reports by both researchers were considered in the next stage, which analyzes if the cause of the bug was introduced in their previous commits. This process requires manual inspection by researchers. In the mean classifying a ticket takes \gema{between 5 and 10 minutes per ticket}, although the amount of time decreases with experience as could be expected.

@@

In this stage we analyzed the 189 tickets classified \grex{Pero no eran 209?} \gema{deberian ser 209 pero me han quedado 20 bugs report por analizar, si tengo tiempo cuando llegue a casa lo hago y anado los numeros} \grex{no se si hace falta, a ver que dice jesus}\gema{Jesus me dijo que no era muy importante a estas alturas} as bug reports and their previous commits.
The possible outcome of the analysis was one of the following three options:

@@

This analysis takes into account that the bug could have been inserted in many lines that belong to many different previous commits, but in fact, not all of them could cause the bug. It may happen that in the previous commit, lines have been copied from further previous commits, comments may have been modified or blank spaces/lines have been introduced. Thus, the cause could be found in a single previous commit, in many or even in none. In the \ref{fig:7} we see that the previous commits are more than one, in this case two \textit{e7be0a988} and \textit{e5296c1da}. The previous commit \textit{e7be0a988} didn't cause the bug because it's a line to know the version of the file. The previous commit \textit{e5296c1da} caused the bug because it inserted a incorrect \textit{break} line. \grex{Todavia no esta del todo claro... yo creo que esto merece un ejemplo - incluso con codigo real, tal y como hemos hecho al principio del articulo. Ademas, nos lo podemos permitir; tenemos espacio.}\gema{perfecto, ahora busco un ejemplo y lo anado abajo}

\begin{figure}[ht]
\centering
\includegraphics[height=2.5cm]{manyprevious.png}
\caption{Previous commits: \textit{e7be0a988}, \textit{e5296c1da} }
\label{fig:7}       % Give a unique label
\end{figure}

@@

As can be seen in Table~\ref{tab:responsability}, from the 308 previous commits, 152 have been considered to be the cause of the bug, whereas 114 were identified as not beign the cause. We were unable to decide in 42 cases, as only lines had been added \ref{fig:8} or due to our limited knowleadge about the code \ref{fig:9}. \grex{Gema, puedes poner un ejemplo de estas dos cosas? Uno de solo added lines y otro de limited knowledge. Si el ejemplo es verdadero, mejor.}\gema{perfecto, ahora busco un ejemplo y lo anado abajo}
We discarded 40 more because they were false positives (\emph{noise}) such as blank lines, changes in comments or even a change in the version of the file.
\begin{figure}[ht]
\centering
\includegraphics[height=2cm]{addedlines.png}
\label{fig:8}       % Give a unique label
\caption{Example of added lines where we are unable to decide which previous commit caused the bug}
%52919
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[height=2cm]{unknown.png}
\label{fig:9}       % Give a unique label
\caption{Example of unknowleadge where we are unable to decide which previous commit caused the bug }
\end{figure}
\begin{table}[htb]
\begin{center}
\begin{tabular}{lrr}
\toprule[0.3mm]
  & \multicolumn{1}{c}{Before} & \multicolumn{1}{c}{After} \\
  & \multicolumn{1}{c}{Deleting Noise} & \multicolumn{1}{c}{Deleting Noise} \\\hline
{Cause} & (152) 44\% & (152) 49\% \\[0ex]
{Not Cause} & (154) 44\% & (114) 37\% \\[0ex]
{Undecided} & (42) 12\% & (42) 14\% \\[0ex]
\bottomrule[0.3mm]
\end{tabular}
\caption{Number of times (and percentage) where the previous commit is the cause, not the cause or could not be classified, before and after deleting noise.}
\label{tab:responsability}
\end{center}
\end{table}

@@

We also studied the distribution of the number of previous commits for each bug. As shown in Table~\ref{tab:secondStage2}, the number of commits that can be considered as previous is 1, followed by 2 commits. \grex{the number of commits that can be considered as previous is 1, followed by 2 commits. No se que queremos decir} \grex{Tenemos que responder a estas dos preguntas todavia: Por que es interesante hacer lo que se comenta en este parrafo? Que utilidad tiene?}\gema{Creo que era interesante saber como se distribuyen los causantes de los bugs en cada projecto y cuantos commits previos presentan los bugs report en cada projecto analizado, ademas cuantos mas commits previos, en ocasiones, es muy dificil saber quien es el responsable por tanto pense que podria haber algun patron en relacion con el numero de commits previos y cuantos de ellos son responsables, pero parece que esto no ocurre con la muestra que hemos analizado.}

\begin{table*}[htb]
\begin{center} {\footnotesize
\begin{tabular}{lccccc}
\toprule[0.3mm]
  & \multicolumn{1}{c}{One previous} & \multicolumn{1}{c}{two previous} & \multicolumn{1}{c}{three previous} & \multicolumn{1}{c}{four previous} & \multicolumn{1}{c}{+five previous}\\
  & \multicolumn{1}{c}{commit} & \multicolumn{1}{c}{commit} & \multicolumn{1}{c}{commit}& \multicolumn{1}{c}{commit}& \multicolumn{1}{c}{commit}\\\hline
\raisebox{1ex}{Neutron} & 11 & 3 & 2 & 2 & 0 \\[0ex]
\raisebox{1ex}{Horizon} & 39 & 8 & 3 & 2 & 4 \\[0ex]
\raisebox{1ex}{Nova} & 44 & 5 & 2 & 4 & 4 \\[0ex]
\raisebox{1ex}{Cinder} & 37 & 9 & 6 & 2 & 2\\[0ex]
\raisebox{1ex}{Total} & 131 & 25 & 13 & 10 & 10\\[0ex]
\bottomrule[0.3mm]
\end{tabular} }
\caption{ Distribution of the number of previous commits per bug report for each project.}
\label{tab:secondStage2}
\end{center}
\end{table*}

@@

In any case, our research shows evidence that assuming that the previous commit is where the cause of a bug can be found does not hold for a not insignificant percentage of bugs. \grex{Tengo que elaborar mas esto, que es el punto fuerte del articulo.}

@@

As any other empirical study, this one presents several threats to its validity, external and internal, that have to be considered and taken into account. In order to allow others to study it in detail, replicate it or even build on top of it, we have set up a replication package including data sources, intermediate data and scripts that can be obtained at the following URL: \gema{Incluir la URL en la que pondre todos los datos y scripts utilizados para los datos.}

@@

\begin{itemize}
  \item We have not considered those tickets where the two researchers showed discordance. 
  \begin{enumerate}
    \item Should they discuss after their analysis to reach a better classification?, Should the tool provide this?
    \item Does the Bug report only the same ticket classified as Bug report for all the researchers?
  \end{enumerate}
    \item We have not taken into account errors that have been classified into \textit{Undecided}, and probably we have lost some \emph{actual} bug reports.
    \item There could be some lax criteria involving the subjective opinion of the researchers.
    \item Although the researchers are experienced programmers, they are not experts in the  OpenStack project, and their inexperience may have influenced the results of the analysis.
    \item We are only using part of the information that the tickets provide, like comments and text. There could be a recognized pattern in the data, unknown at first sight, that involves other parts of the information.
    \item We have used a random script to extract the tickets from Launchpad that have been reported during 2015. There could be unintended bias of the data, because many reasons, as for instnace the phase of the project.
    \item In some cases, researchers may have classified the previous commit as the cause of the bug, even if this may not be the case (see discussion on additional conditions in if statements).
\end{itemize}

@@

With our results we can identify which ones are real changes that introduced the bug, and this could be useful to improve the accuracy of those tools developed to prevent bugs. Also, the software developers stand to benefit from identifying where the bug was inserted, improving their methodology.
%How to plan to apply the knowledge in a way that can benefit software engineers.
%What are the anticipated contributions of the work?  How will you evaluate them to demonstrate usefulness?

A final field in our future work concerns the full automation of the methodology could developer an automatic classifier base on the idea that not all the previous commit injected the bug. Another interesting investigation could perform the same empirical study in a project with a community less active, to can prove if our idea is fulfill in other projects.\grex{Puedes aclarar esto, porfa? En castellano si quieres}
\gema{Si, la idea es realizar el mismo estudio pero con otro projecto que no sea OpenStack, porque en realidad OpenStack esta evolucionando continuamente. Y para poder demostrar que en realidad la idea se cumple en todos los projectos, deberiamos analizar un projecto que sea mas 'estable' y en el que la API no se este cambiando continuamente }
